Gozde and Liora Replication Code

### 2015-03-16 Gozde-Liora-Replication-Code.Rmd
replication submission

@@

```{r}
library(Zelig)
library(lme4)
library(MASS)
library(nlme)
library(plm)
library(xtable)
library(foreign)
library(dplyr)
library(broom)
library(ggplot2)
library(xtable)
library(stargazer)

@@

# Cleaning up data
data <- rename(data, ext_vote = rexvote)
data <- rename(data, age_1 = age1) # Age (16-25)
data <- rename(data, age_2 = age2) # Age 
data <- rename(data, age_4 = age4) # Age (65- )
data <- rename(data, edu_1 = mye1) 
data <- rename(data, edu_2 = mye2)
data <- rename(data, age_2 = age2)
data <- rename(data, farmer_own = farmerown)
data <- rename(data, left_right_scale = zlrs)
data <- rename(data, eu_neg = euschlecht) # Negative evaluation of EU membership of one's own country
data <- rename(data, z_dem_satis = zsatisdmo) 
data <- rename(data, dispro_elect = disp) 
data <- rename(data, federalism = lfed1)
data <- rename(data, z_asylumseeker = zasylumseekers)
data <- rename(data, z_unemp = zsur) 
data <- rename(data, z_replacement = zreplacementrate) 
data <- rename(data, max_er = rmax)

@@

# Create centered variables for Saliance and Variance based on grand mean centering according to author

```{r}
data <- data  %>% 
  mutate(salience_mean_c = (salienzmean - 3.84568)) %>% 
  mutate(var_c = rvar - 21.75423)
```

# Make male variable numeric

```{r}
data <- data %>%
  mutate(male = as.numeric(male))
```

# Create a dummy variable for country (transform from character to factor variable)

```{r}
data <- data %>%
  mutate(country = factor(sortcountry, labels = c("AT", "BE","DE-E","DE-W","DK","ES","FI","FR","GR","IT","LU","NL","NO","PT","SE")))
```

# Plot Data for exploration

```{r}
ggplot(data, aes(x = z_unemp, y = ext_vote)) + geom_point() + geom_smooth(method = "glm", family = "binomial", se = FALSE) # Relationship w/ standardized unemployment rate

ggplot(data, aes(x = z_asylumseeker, y = ext_vote, color = country)) + geom_point() + geom_smooth(method = "glm", family = "binomial", se = FALSE) # Relationship w/ immigration

ggplot(data, aes(x = z_replacement, y = ext_vote, color = country)) + geom_point() + geom_smooth(method = lm, se = FALSE) #Relationship w/ welfare (=unemp benefit)
```

@@

```{r}
model <- glmmPQL(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + left_right_scale + eu_neg + z_dem_satis + dispro_elect + federalism + z_asylumseeker + z_unemp + z_asylumseeker:z_unemp + z_replacement + z_replacement:z_unemp + z_replacement:z_asylumseeker + max_er + salience_mean_c + var_c + var_c:salience_mean_c + country - 1, random = ~1|kontext, family = binomial(link = "logit"), data = data, verbose = TRUE)
summary(model)

model1 <- glmmPQL(rexvote ~ male + age1 + age2 + age4 + mye1 + mye2 + farmerown + worker + retired + unemployed + zlrs + euschlecht + zsatisdmo + disp + lfed1 + zasylumseekers + zsur + zasylumseekers:zsur + zreplacementrate + zreplacementrate:zsur + zreplacementrate:zasylumseekers + rmax + salienzmean.c + rvar.c + rvar.c:salienzmean.c + country - 1, random=~1|kontext, family=binomial(link="logit"), data=data, verbose=TRUE)
summary(model1)

# Model 2 based on glm logit - comparison

```{r}
model2 <- glm(rexvote ~ male + age1 + age2 + age4 + mye1 + mye2 + farmerown + worker + retired + unemployed + zlrs + euschlecht + zsatisdmo + disp + lfed1 + zasylumseekers + zsur + zasylumseekers:zsur + zreplacementrate + zreplacementrate:zsur + zreplacementrate:zasylumseekers + rmax + salienzmean.c + rvar.c + rvar.c:salienzmean.c + country - 1, family=binomial(link="logit"), data=data, verbose=TRUE)
summary(model2)

#Table 1

@@

coefficient <- as.vector(model$coef$fixed) 
n <- c(174452, 267)

stand_dev <- as.vector(sqrt(diag(model$varFix)))

var_names <- c("Male","18–29 years","30–45 years",">65 years","Education: middle/high","Education: university","Petty Bourgeoisie","Worker","Pensioner","Unemployed","Left-Right","Dissatisfied: EU","Dissatisfied: Democracy","Disproportionality","Decentralization","Asylumseekers", "Unemployment", "Asylumseekers × Unemployment", "Toughness", "Salience", "Variance", "AT", "BE", "DE-E", "DE-W", "DK", "ES", "FI", "FR", "GR", "IT", "NL", "NO", "PT", "SE", "Asylumseekers x Unemployment", "Unemployment benefits x Unemployment", "Unemployment benefits x Asylumseekers", "Variance x Salience")

@@

##Label matrix.

@@

# Create table using xtable

x <- xtable(table_matrix, digits = 3, caption = "Logistic multilevel model. PQL2 estimates, model-based standard errors in parentheses. MI results based on 11 separate imputations.")
print.xtable(x, type = "html", file="tablepractice.html")

# Create Table 1 using stargaze

stargazer(table_matrix, title = "TABLE 1 Support for the Extreme Right: Sociodemographics, Attitudes, Country Effects, and Contextual Variables", align = TRUE, table.layout = "t", notes = "Logistic multilevel model. PQL2 estimates and model-based standard errors.", style = "ajps", out = "Table 1.html", no.space = TRUE)

@@

```{r}
library(msm)
asy <- seq(from = -.98, to = 4.46, by = .01) # vector with possible values for asylumseekers
variable <- asy
slopes <- model$coef$fixed[17] + model$coef$fixed[36] * asy # Estimated slopes based on estimated coefficients of unemployment (17) and unemployment x immigration (36)
estmean <- model$coef$fixed # Estimated means for each variable 
var <- vcov(model)
se <- rep(NA, length(asy))

@@

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Asylumseekers"
ytitle <- "Effect: Unemployment"
main <- "Figure 1: The Conditional Effects of Unemployment and Immigration"
graph1 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle, title = main)

@@

@@

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Unemployment"
ytitle <- "Effect: Asylum seekers"
main <- "The Conditional Effects of Unemployment and Immigration"
graph2 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle)

# Figure 1 with ggplot
library(grid)
library(gridExtra)
grid.arrange(graph1, graph2, ncol = 1)

@@

# Figure 2, graph 1 (Conditonal Effects of Salience and Variance)

salience <- seq(from = -3.108026, to = 21.872906, by=.1) # vector with possible values for salience
variable <- salience
slopes <- model$coef$fixed[21] + model$coef$fixed[39] * variable # Estimated slopes based on estimated coefficients of salience (20) and salience x variance (39)
estmean <- model$coef$fixed # Estimated means for each variable 
var <- vcov(model)
se <- rep(NA, length(variable))

@@

# Creating the same graph with ggplot - much better looking.
library(ggplot2)
variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Salience"
ytitle <- "Effect: Variance"
main <- "The Conditional Effects of Salience and Variance"
graph3 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle, title = main)
```

>>>>

Note: Please refer to saved files in our repository for images from the article. Also please refer to the table and figures (Table 1, Figure 1, 2) in the repository. Image scales are rearranged in saved versions. Our regression values are slightly different from the authors as his stata code used a slightly different method than our function (glmmPQL) to calculate Penalized Quasi-likehood. Finally, we were not able to reproduce the last two figures that showed the joint impact of more than 2 variables.

#The paper
Kai Arzheimer, "Contextual Factors and the Extreme Right Vote in Western Europe, 1980-2002"25548117.pdf
This article attempts to answer the “twin question of why the extreme right support is so unstable within many countries over time, and why these parties are so weak in many West European countries.” Authors conduct a multi-level analysis, combining data on macro-level contextual variables (unemployment, immigration, and welfare benefits) with individual-level variables (sociodemographic factors and attitudes).

#The data
Individual-level variables come from Eurobarometer surveys (1980-2002), while contextual variables are gathered from a number of different data sets produced by the OECD, UNHCR etc. The merged data and the Stata code are available through the author’s dataverse at: http://hdl.handle.net/1902.1/12092.

#The model 
The article uses a logistic multi-level model via Penalized Quasi-Likelihood. The dependent variable is vote intention for an extreme right party, which is assumed to be binomially distributed.

@@

library(MASS)
library(nlme)
library(xtable)
library(foreign)
library(dplyr)
library(broom)
library(ggplot2)
library(xtable)
library(stargazer)
library(grid)
library(gridExtra)

@@

# Cleaning up data
data <- rename(data, ext_vote = rexvote)
data <- rename(data, age_1 = age1) # (18–29 years)
data <- rename(data, age_2 = age2) # (30–45 years)
data <- rename(data, age_4 = age4) # (>65 years)
data <- rename(data, edu_1 = mye1) # Education: middle/high
data <- rename(data, edu_2 = mye2) # Education: university
data <- rename(data, farmer_own = farmerown) 
data <- rename(data, left_right_scale = zlrs) # Left-Right Scale
data <- rename(data, eu_neg = euschlecht) # Negative evaluation of EU membership of one's own country
data <- rename(data, z_dem_satis = zsatisdmo) # Dissatisfied: Democracy
data <- rename(data, dispro_elect = disp) # Disproportionality
data <- rename(data, federalism = lfed1) # Decentralization
data <- rename(data, z_asylumseeker = zasylumseekers) 
data <- rename(data, z_unemp = zsur) # Unemployment rate
data <- rename(data, z_replacement = zreplacementrate) # Unemp benefits
data <- rename(data, max_er = rmax) # Maximum toughness in extreme right discourse

@@

# Preparing data for analysis

```{r}
# Create centered variables for Saliance and Variance based on grand mean centering according to author

data <- data  %>% 
  mutate(salience_mean_c = (salienzmean - 3.84568)) %>% 
  mutate(var_c = rvar - 21.75423)

# Make male variable numeric
data <- data %>%
  mutate(male = as.numeric(male))

# Create a dummy variable for country (transform from character to factor variable)
data <- data %>%
  mutate(country = factor(sortcountry, labels = c("AT", "BE","DE-E","DE-W","DK","ES","FI","FR","GR","IT","LU","NL","NO","PT","SE")))
```

@@

```{r}
model <- glmmPQL(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + left_right_scale + eu_neg + z_dem_satis + dispro_elect + federalism + z_asylumseeker + z_unemp + z_asylumseeker:z_unemp + z_replacement + z_replacement:z_unemp + z_replacement:z_asylumseeker + max_er + salience_mean_c + var_c + var_c:salience_mean_c + country - 1, random = ~1|kontext, family = binomial(link = "logit"), data = data, verbose = TRUE)
# summary(model) # Output too long
```

# Table 1

@@

coefficient <- as.vector(model$coef$fixed) 
n <- c(174452, 267)
stand_dev <- as.vector(sqrt(diag(model$varFix)))

@@

# Label matrix.

@@

# Create Table 1 with stargaze

table_1 <- stargazer(table_matrix, title = "TABLE 1 Support for the Extreme Right: Sociodemographics, Attitudes, Country Effects, and Contextual Variables", align = TRUE, table.layout = "t", notes = "Logistic multilevel model. PQL2 estimates and model-based standard errors.", style = "ajps", out = "Table 1.html", no.space = TRUE)

@@

```{r}
library(msm)
asy <- seq(from = -.98, to = 4.46, by = .01) # vector with possible values for asylumseekers
variable <- asy # to standardize the code
slopes <- model$coef$fixed[17] + model$coef$fixed[36] * asy # Estimated slopes based on estimated coefficients of unemployment (17) and unemployment x immigration (36)
estmean <- model$coef$fixed # Estimated means for each variable 
var <- vcov(model) # Estimated variances
se <- rep(NA, length(asy))

@@

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Asylumseekers"
ytitle <- "Effect: Unemployment"
main <- "Figure 1: The Conditional Effects of Unemployment and Immigration"
graph_1 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle, title = main) +
  coord_fixed(ratio = 8)

@@

# Creating second part of the figure in base R

@@

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Unemployment"
ytitle <- "Effect: Asylum seekers"
main <- "The Conditional Effects of Unemployment and Immigration"
graph_2 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle) +
  coord_fixed(ratio = 12)

# Figure 1 with ggplot

grid.arrange(graph_1, graph_2, ncol = 1)

# Plot separately - better scaled
graph_1
graph_2

@@

# Figure 2 (Conditonal Effects of Salience and Variance)

```{r}
# Graph 1
salience <- seq(from = -3.108026, to = 21.872906, by=.1) # vector with possible values for salience
variable <- salience
slopes <- model$coef$fixed[21] + model$coef$fixed[39] * variable # Estimated slopes based on estimated coefficients of salience (21) and salience x variance (39)
estmean <- model$coef$fixed # Estimated means for each variable 
var <- vcov(model)
se <- rep(NA, length(variable))

@@

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Salience"
ytitle <- "Effect: Variance"
main <- "The Conditional Effects of Salience and Variance"
graph_3 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle, title = main) +
  coord_fixed(ratio = 400)

# Second part of Figure 2

variance <- seq(from = -21.71426, to = 174.34497, by=10) # vector with possible values for variance
variable <- variance
slopes <- model$coef$fixed[20] + model$coef$fixed[39] * variable # Estimated slopes based on estimated coefficients of salience (20) and salience x variance (39)
se <- rep(NA, length(variable))
for (i in 1:length(variable)) {
  j <- variable[i]
  se[i] <- deltamethod (~ (x20) + (x39) * j, estmean, var)
}
# Create confidence intervals and combine all values in matrix.
upper <- slopes + 1.96 * se
lower <- slopes - 1.96 * se
variable.data <- cbind(variable, slopes, upper, lower)

# Creating the second graph in base R 
par(mfrow = c(1,1))
plot(variable, slopes, type = "l", lty = 1, ylim = c(-.05, .2), xlim = c(-50, 150), ylab = "Effect: Saliance", xlab = "Variance")
points(variable, upper, type = "l", lty = 2)
points(variable, lower, type = "l", lty = 2)
points(variable, rep(0, length(variable)), type = "l", col = "black")

variable.data <- data.frame(variable.data) # Turn matrix into data frame so that ggplot can read it.
xtitle <- "Variance"
ytitle <- "Effect: Salience"
main <- "The Conditional Effects of Salience and Variance"
graph_4 <- ggplot(variable.data, aes(x = variable, y = slopes)) + 
  geom_line(aes(y = slopes), color = "blue") + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  geom_hline(yintercept = 0, size = 0.4, color = "black") +
  labs(x = xtitle, y = ytitle, title = main) +
  coord_fixed(ratio = 500)

grid.arrange(graph_3, graph_4, ncol = 1)
# Better scaled versions - not combined
graph_3
graph_4
```
